---
title: "PasswordGen R package"
author: "Marius Pascariu"
date: "`r Sys.Date()`"
output: rmarkdown::html_vignette
# output: md_document
vignette: >
  %\VignetteIndexEntry{Vignette --- PasswordGen R package}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

Sometimes we need not only secure passwords but also passwords easy to remember.
PasswordGen is an **R** package that helps you generates passwords and passphrases 
easy to remember with an algorithm based on the dictionary method. 
 
The default ```password()``` function generates random passwords using words from 
[The Project Gutenberg EBook of Webster's Unabridged Dictionary, 
by Various](http://www.gutenberg.org/ebooks/29765). In total 93077 words are used.
However, you can use the ```ExtractWords()``` function to extract all the words from 
any book and use them in the ```password()``` generator. 
Check the examples provided in the help files of the package.   


### Installation

1. Make sure you have the most recent version of R
2. Run the following code in your R console 

```{r, eval=FALSE}
devtools::install_github("mpascariu/PasswordGen")
```

If you do not have __devtools__, you can install it from CRAN, by typing `install.packages("devtools")`.

Alternatively, clone or download the repository to a local directory and install it by running `install.packages(<local_dir>, repos = NULL, type = "source")`.


### Help
All functions are documented in the standard way, which means that 
once you load the package using ```library(PasswordGen)```
you can just type ```?password``` to see the help file. 




### Example 1

Generate passwords and passphrases using \code{password} function.  

```{r}
library(PasswordGen)
password()
```
The default version of the function is generating a string that containts three words (lower and upper cases) plus some random symbols and numbers. As mentioned above the words are borrowed from Webster's Unabridged Dictionary.

### Example 2
We may want to change increase or decrease the complexity of our passwords. We can do this by playing with the arguments of the \code{passwords} function.

Let's generate a passphrase composed out of 5 words plus a random number after each word, no symbols, and separate them using a dash line.
```{r}
password(no_words = 5,     # 5 word password
         numbers = TRUE,   # with numbers
         symbols = FALSE,  # without symbols
         sep = '-') # with a space between the words
```

### Example 3
Generate a password using words from my favorite book.

Using the \code{gutenbergr} R package we will download a book and
extract words from it. "The Adventures of Sherlock Holmes" by
Doyle, Arthur Conan is a good choice.


```{r}
library(gutenbergr)
library(stringr)
# look up book and check the id on Gutenberg Project
info <- gutenberg_works(str_detect(title, "The Adventures of Sherlock Holmes"))
info$gutenberg_id # this should be 1661

# download the book
my_book <- gutenberg_download(gutenberg_id = 1661, meta_fields = "title",
                             mirror = "http://gutenberg.pglaf.org")
```

Now we can extract the words using the \code{ExtractWords} function.

```{r}
words <- ExtractWords(data.r = my_book$text)
words
# Note, this book has significantly less unique words than Webster's Unabridged Dictionary. More words to choose from is better.
```

And generate the password:

```{r}
password(no_words = 5, data = words)
```

### Example 4
Extract words from other language than English and generate passwords. 
Mihai Eminescu was a Romantic poet, often regarded as the most famous and influential Romanian poet. Let's check his words.

```{r}
# check the id of the book on Gutenberg Project
info <- gutenberg_works(languages = 'ro')

# download the book
my_book <- gutenberg_download(gutenberg_id = 35323, meta_fields = "title",
                             mirror = "http://gutenberg.pglaf.org")

# Extract words
words <- ExtractWords(data.r = my_book$text)

password(no_words = 4, data = words,
        numbers = TRUE, symbols = TRUE, sep = "_")

```
Nice Romanian password!

